---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# statsdata

<!-- badges: start -->
[![R-CMD-check](https://github.com/PreTeXtBooks/statsdata/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/PreTeXtBooks/statsdata/actions/workflows/R-CMD-check.yaml)
[![Lifecycle: stable](https://img.shields.io/badge/lifecycle-stable-brightgreen.svg)](https://lifecycle.r-lib.org/articles/stages.html#stable)
<!-- badges: end -->

`statsdata` provides the datasets used in the PreTeXtBooks open statistics 
textbook series. It is intended for students working through the books, 
and for instructors who want to reproduce or adapt the examples and 
exercises.

## Installation

Install the package from GitHub with:

```{r, eval = FALSE}
# install.packages("pak")
pak::pak("IDEMSInternational/statsdata")
```

## Datasets

| Dataset | Description | Source |
|---|---|---|
| `kenya_flamingos` | Simulated morphological measurements for Greater and Lesser Flamingos at three Rift Valley lakes | Simulated for teaching purposes |
| `kenya_accidents_2016` | Fatal road accident records from Kenya, April–June 2016 | [Humanitarian Data Exchange](https://data.humdata.org/dataset/kenya-road-accidents-database) |
| `kenya_accidents_2017` | Fatal road accident records from Kenya, 2017 | [Humanitarian Data Exchange](https://data.humdata.org/dataset/kenya-road-accidents-database) 

Raw xlsx files for use with other software are available in the 
[`data-raw/`](https://github.com/PreTeXtBooks/statsdata/tree/main/data-raw) 
folder.

## Usage

```{r example}
library(statsdata)

head(kenya_flamingos)
```

```{r species-summary}
library(dplyr)

kenya_flamingos %>%
  group_by(species) %>%
  summarise(
    n             = n(),
    mean_mass_g   = round(mean(body_mass_g, na.rm = TRUE)),
    mean_bill_mm  = round(mean(bill_length_mm, na.rm = TRUE), 1)
  )
```

## Notes on simulated data

The `kenya_flamingos` dataset was simulated for teaching purposes. 
The measurements are realistic for the two species at these lakes.
It has been simulated to be like the `iris` or `penguins` data set,
but more contextually relevant for Kenya.
It should not be used as a source ofempirical claims about flamingo
populations in Kenya.

## Licence
Individual datasets carry their own licences as documented in the help pages 
(`?kenya_flamingos`, `?kenya_accidents_2016`). The Kenya road 
accidents data is released under the Humanitarian Data Exchange 
[HDX Open Data Licence](https://data.humdata.org/faqs/licenses).
